\input{chapters/extras/concept_lattice_to_mlb.tex}
\subsection{\acl{mlb} Constraints}
\label{sec:foundations_mlb}

The purpose of clustering is to structure data without any class labels.
This is especially useful in real world applications since data often lacks labels and the annotation process is expensive.
% Partitional clustering groups data points into clusters based on a similarity measure~\cite{bade_hierarchical_2014}.
% With hard clustering, each data point belongs to only one cluster, while soft clustering allows membership to multiple clusters~\cite{schutz_comprehensive_2023}.
% 
Clustering algorithms group data points based on different criteria into clusters.
The overall goal is to minimize the intra-cluster distances and maximize the inter-cluster distances.
Prior knowledge about the dataset can be modelled as constraints. %~\cite{bade_hierarchical_2014}.
The simplest constraints are pairwise \textit{must-link} and \textit{cannot-link} constraints.
% Pairwise constraints can be either instance-based or metric-based.
% Instance-based constraints directly the clustering algorithm by enforcing constraints on their instance pairs.
% They can be hard or soft, allowing different degrees of constraint violation.
% Metric-based constraints map the constraints to the distance metric or similarity measure of the clustering algorithm.
Pairwise constraints are limited to partitional clustering.
% \subsubsection{Hierarchical clustering}
Opposed to partitional clustering, hierarchical clustering groups the data on different levels of granularity~\cite{bade_hierarchical_2014}.
% Hence, partitional clustering is a special case of hierarchical clustering with one level of granularity.
So-called \ac{mlb} constraints introduce prior knowledge into hierarchical clustering.
A \ac{mlb} constraint
\begin{eqnarray}
    MLB_{x y z} = (d_x, d_y, d_z)
    \label{eq:mlb_xyz}
\end{eqnarray}
specifies an agglomerating order on three data points $d_x, d_y, d_z$.
Symmetry applies for the first two elements.
For the \ac{mlb} constraint in \eqref{eq:mlb_xyz} both \autoref{def:mlb_implication} and \autoref{def:mlb_counterexample} must hold true.
This means that the must-link constraint between the first and second item is stronger than that between the first and third item.
We illustrated the resulting \ac{mlb} constraints on a small toy example in \autoref{fig:hierarchy_to_mlb}.
\begin{definition}[\ac{mlb} Implication]
    \label{def:mlb_implication}\\
    Given \ac{mlb} constraint $MLB_{x y z}$:
    $\forall c \in H: d_x \in c \wedge d_z \in c \rightarrow d_y \in c$
\end{definition}
\begin{definition}[\ac{mlb} Counter Example]
    \label{def:mlb_counterexample}\\
    Given \ac{mlb} constraint $MLB_{x y z}$:
    $\exists c \in H: d_x \in c \wedge d_y \in c \wedge d_z \notin c$
\end{definition}

\hierarchyToMLB{}

% A \ac{mlb} constraint for hierarchical clustering can be transformed into multiple pairwise constraints for partitional clustering by adding one must-link constraint between $d_x$ and $d_y$ and two cannot-links between $d_x, d_z$ and $d_y, d_z$~\cite{bade_hierarchical_2014}.

% \subsubsection{Concept hierarchies} 
% Concept hierarchies structure information into hierarchies and facilitate working with knowledge-bases~\cite{cimiano_learning_2005}.
