\begin{abstract}
\ac{fca} enables the generation of concept hierarchies without requiring an expert ontology.
This raises the question of whether the resulting hierarchy aligns with existing sources of knowledge. 
To guide the construction of coherent concept hierarchies, prior work in constrained clustering introduces the use of hierarchical \ac{mlb} constraints.

In this work, we introduce a method for comparing and integrating knowledge derived from \acs{mlb} constraints with an additional hierarchical knowledge source. 
To evaluate the benefits of our approach, we conduct experiments on the BankSearch dataset and explore the universe of a closure operator provided by the \acs{mlb} constraints.
These constraints are derived from a gold-label topic hierarchy and subsequently mapped onto the corresponding document corpus.
As our supplementary hierarchical knowledge source, we employ a \acl{lda} topic model, which constructs a document-topic context through statistical inference over the same corpus.

We mathematically prove that MLB constraints induce a valid closure operator, providing the theoretical foundation for integrating hierarchical expert knowledge into FCA. 
We then analyze the structural differences between the concept lattices guided by the two knowledge sources. 
The influence of both knowledge bases is assessed qualitatively through manual review and quantitatively by analyzing their intersection for different levels of required coherence.
We find distinct structural variations between the two concept lattices, reflected in their \textit{consistency} and \textit{coherence} lattice intersection.
Our findings suggest a significant divergence between expert-defined hierarchies and data-driven topics, highlighting the necessity of relaxed coherence measures for practical knowledge integration.

\keywords{\acl{fca} \and Constraints \and Topic Modeling.}
\end{abstract}