\section{Introduction}
\label{sec:introduction}

% WORDNET: manual creation cf. le2019inferring, hearst1992hyponyms; hyponym = is-a relationship cf. hearst1992hyponyms
Concept hierarchies form the backbone of expert ontologies such as WORDNET~\cite{miller1990wordnet,fellbaum1998wordnet}.
They consist of systems of hyponym relationships~\cite{le2019inferring} organized under a single top-level node.
By structuring information into categories and enabling the formulation of rules and relations, concept hierarchies facilitate the retrieval and reuse of knowledge bases~\cite{cimiano_learning_2005,yang2012personalized}.
However, the manual construction of high-quality domain-specific ontologies is costly and typically reflects only a snapshot of evolving knowledge~\cite{li2018unsupervised}.

Automatic concept hierarchy generation offers a scalable alternative and enables the construction of hierarchies in domains not covered by existing expert ontologies.
Such approaches commonly rely on the identification of explicit textual patterns~\cite{le2019inferring}, for example Hearst patterns~\cite{hearst1992hyponyms}, or on \ac{nlp} techniques that extract latent information from text and represent it in the form of binary object-attribute matrices, which can then be analyzed using \ac{fca}~\cite{cimiano_learning_2005}.
In contrast to traditional similarity-based approaches, \ac{fca}-based concept hierarchy generation can be viewed as a form of conceptual clustering that yields intensional descriptions of abstract concepts~\cite{cimiano_learning_2005}.

Although manually curated ontologies are often regarded as the gold standard, automatic generation is desirable to reduce time and cost.
However, automatically generated hierarchies may conflict with established expert ontologies, potentially undermining trust in one or both knowledge sources.
Ensuring that generated hierarchies remain sufficiently consistent with existing hierarchical background knowledge therefore requires mechanisms to guide and constrain the generation process.

Previous work on guidance mechanisms has explored human involvement~\cite{yang2012personalized} as well as the use of \acp{llm}~\cite{kommineni2024from}. 
Technically, guidance has been implemented by restricting the construction of concept lattices through formal constraints~\cite{belohlavek2006formal,belohlavek2009formal} or by incorporating attribute weights~\cite{belohlavek2011selecting}.
Furthermore, Bade and Nürnberger~\cite{bade_personalized_2006,bade_hierarchical_2014} proposed the use of hierarchical constraints to generate concept hierarchies within similarity spaces derived from text.

% Besides presentation purposes, \ac{fca} has been employed to identify potentially missing thesaurus entries in a specific expert ontology~\cite{zheng2020lexical,hao2021leveraging}. 

To date, the systematic integration of hierarchical background knowledge into \ac{fca}-based concept hierarchy generation through \ac{mlb} constraints has received limited attention.
We address this gap by incorporating \ac{mlb} constraints into the \ac{fca} framework to guide the construction of concept hierarchies.
In addition, we present a case study in which \ac{fca}-based hierarchies generated under \ac{mlb} constraints are compared to an independent knowledge source derived from corpus-specific statistical topic modeling.

% structure of the paper
The remainder of this paper is structured as follows: In Section~\ref{sec:related_work}, we review related work on concept hierarchy generation and the use of constraints in \ac{fca}. 
In Section~\ref{sec:foundations}, we provide the necessary background on \ac{fca} and \ac{mlb} constraints. 
In Section~\ref{sec:method}, we describe our proposed method for integrating hierarchical background knowledge in \ac{fca} via \ac{mlb} constraints. 
In Section~\ref{sec:case_study}, we present a case study comparing \ac{fca}-based concept hierarchies generated with \ac{mlb} constraints to a second knowledge base based on corpus-specific statistical topic modeling. 
Finally, in Section~\ref{sec:future_work}, we discuss potential future work and conclude the paper.

%\textcolor{red}{NOTE: Their conclusion includes future work on using humans for ontological relation validation (Attr. Explor.???). They also compare generated ontologies with human-modeled ontologies via Precision, Recall, F-measure...}

% Usually you
%Since this approach relies solely on syntactic structure present in the text corpus, the available knowledge is very limited.
%Hierarchical constraints can be thought of a tool to model certain degrees of certitude with respect to the information's validity.
