\section{Related Work}
\label{sec:related_work}

Clustering algorithms group data points based on different criteria into clusters.
Prior knowledge about the dataset can be modelled as constraints.
The simplest constraints are pairwise \textit{must-link} and \textit{cannot-link} constraints, which require objects to be in the same or different clusters, respectively.
While conventional clustering groups data points into clusters on a single level of granularity, hierarchical clustering groups the data on different levels of granularity~\cite{bade_hierarchical_2014}.
So-called \ac{mlb} constraints introduce prior knowledge into hierarchical clustering.

In hierarchical clustering, two main approaches exist: partitional and agglomerative hierarchical clustering.
While partitional hierarchical clustering build a hierarchy by recursively partitioning the data (i.e., top to bottom), agglomerative hierarchical clustering methods build a hierarchy by iteratively merging clusters (i.e., bottom to top)~\cite{zhao2002evaluation}.

Constrained partitional clustering methods such as constrained $k$-means clustering~\cite{wagstaff2001constrained} assigns documents to the closest cluster prototype that does not violate any constraints \cite{wagstaff2001constrained}.
Pairwise constraints are inherently ambiguous in hierarchical settings because objects may be together at one level of the hierarchy but separated at another.

Bade et al.~\cite{bade2006personalized,bade_creating_2008} find pairwise constraints to be insufficient for hierarchical clustering, and introduce \ac{mlb} constraints into \ac{hac}.
These \ac{mlb} constraints encode the order in which objects should be merged during agglomeration, effectively capturing relative hierarchical relationships rather than flat equivalence conditions~\cite{bade2006personalized,bade_creating_2008,bade_hierarchical_2014}. % or: grouping data on different granularities
% The authors formally analyze the properties of \ac{mlb} constraints, including symmetry and transitivity, and propose an instance-based extension of \ac{hac} that enforces these constraints during the merge process. 
% Empirical results on text clustering tasks show that incorporating \ac{mlb} constraints provides effective structural bias, improving the recovery of meaningful hierarchies while avoiding the level-selection problems associated with adapting pairwise constraints to hierarchical clustering.
% The \ac{hac} algorithm iteratively merges the two most similar clusters into one. The \ac{ihac} extends the \ac{hac} algorithm by not only considering similarity but also constraint violations~\cite{bade_hierarchical_2014}.
% \textcolor{red}{TODO: How to get constraint base \& representation of it.}
While Bade et al.~\cite{bade2006personalized} use \ac{mlb} constraint violation to alter the similarity measure during the merge process, Zhao et al.~\cite{zhao2010hierarchical} propose a method also based on trenary order constraints, denoted ordering-constraints, that delays merging two clusters until violated constraints are resolved.
Zheng et al.~\cite{zheng2011semi} porpose two matrices-based approaches for incorporating \ac{mlb} constraints into hierarchical clustering. %, which are based on a distance minimization and transformation XY, respectively.

MacLellan et al.~\cite{maclellan2016trestle} incrementally builds concept hierarchies.


The generation of concept hierarchies by incorporating restrictive information is currently described as a hierarchical text classification task, which represents a finer structuring task than clustering. % maybe claim from \cite{li2019hiercon} related work??? abstract risch2020hierarchical? abstract: yu2022constrained yes
Such restrictive hierarchical classification approaches are usually based on more probabilistic or neural architectures. Gopal and Yang (2013)~\cite{gopal2013recursive}, for example, proposed using recursive
regularization constraints for classifiers at different hierarchical levels. They showed that large concept hierarchies with 600\,000 classes can be learned efficiently provided that sufficient training data is available.

Li et al.~\cite{li2019hiercon} embedded both documents and hierarchical categories into a common space by representing them as distribution of words appearing in the document corpus.
The similarity between document and category representations is then used to label documents with their most similar categories. 
Since they work with hierarchical categories, labeling documents with one or more categories also implicitly can be considered as a hierarchical clustering of documents.
% wie kommt LDA da jetzt rein?
% Following the concept distribution assumption for text documents—most famously implemented by \ac{lda}~\cite{blei2003lda}—it has been observed that technical documents are typically
% organized as a distribution over concepts that contain discriminative information about their topics. It has been proposed to integrate constraints for topic-based concept hierarchies~\cite{li2019hiercon}. % concepts are basically words from the document corpus, using the term is really confusing here!
% Such topic-based concept hierarchies project both categories and documents into a common space, whereby the similarity between terms can be taken into account. Li et al. (2019)~\cite{li2019hiercon} proposed
% a concept-based framework for generating topic-based concept hierarchies for technical documents, which uses only a small set of documents labeled and organized as a concept hierarchy as input.

\textcolor{orange}{(SST): Please ensure that \ac{lda} is referenced in later sections approriately. You are not credible if you use a new hierarchical \ac{lda} approach but do not reference \ac{lda} appropriately. Klara: Not sure if we need reference \ac{lda} here, we have a chapter for that.}

Risch et al.~\cite{risch2020hierarchical} use attention-based neural sequence-to-sequence models to generate hierarchical labels (i.e., a sequence of labels starting with broad ones and ending with fine-grained ones) for documents.
Similarly, Yu et al.~\cite{yu2022constrained} propose a encoder-decoder architecture trained to assign multiple hierarchical labels to documents constraining assigned labels to be hyponyms of each other. 
Zhang et al.~\cite{zhang2025teleclass} employ \acp{llm} to enrich an existing labels and label taxonomy (i.e., hierarchical structure of categories) with indicative terms and subsequently match embedded documents and categories.
% For patents, Risch et al. (2020)~\cite{risch2020hierarchical} formulate a sequence generation task that additionally uses semantic information about parent-child relationships as constraints during the
% training process, whereby labels are only created if they are consistent within a hierarchy. Currently, the problem of “label consistency” is further addressed by other approaches based on constrained
% sequence-to-tree generation tasks~cited in~\cite{yu2022constrained,zhang2025teleclass}.

\todo{TODO (SST): Why not think about in a first step about such modern approaches? What could be an advantage of a more simple approach?}

\todo{TODO: topic models (to doc-topic incidence) -> SST: There is already reference to Blei (2003) probably that is enough in a first step. The neural approaches are more interesting}
% \input{chapters/03_related_work/doc_topic.tex}

\todo{TODO (SST): The following references are more like an internal discussion not really related work. The references should been make at later sections if approriate.}

% Attribute Exploration with contradicting experts
\cite{felde2022attribute}: Attribute Exploration with contradicting experts

% constrained FCA
\cite{hidayat_constrained_fca_2024} propose an approach for computing a non-redundant implicational base of a formal context with constraints using SAT solvers.
Their constraints encode certain attributes that are mutually exclusive.

\cite{mao_constrained_fca_2017} allow the user to specfiy realtive attribute importance as background knowledge in \ac{fca}.
Their approach reduces the number of fromal concepts extracted from the input data by focusing on those that involve more important attributes.