\section{Foundations}
\label{sec:foundations}

\subsubsection{Constraints}
Constraints model prior knowledge about an optimum~\cite{bade_hierarchical_2014}.


\subsubsection{Hierarchical clustering}
The purpose of clustering is to structure data without any class labels.
This is especially useful in real world applications since data often lacks labels and the annotation process is expensive.
Partitional clustering groups data points into clusters based on a similarity measure.
Hierarchical clustering adds another level of usefulness by grouping the data on different levels of granularity~\cite{bade_hierarchical_2014}.
Hence, partitional clustering is a special case of hierarchical clustering with one level of granularity.

Clustering algorithms group data points based on different criteria into clusters.
The overall goal is to minimize the inner-cluster distances and maximize the intra-cluster distances.
Prior knowledge about the dataset can be modelled as constraints.
The simplest constraints are pairwise \textit{must-link} and \textit{cannot-link} constraints.
Pairwise constraints can be either instance-based or metric-based.
Instance-based constraints directly the clustering algorithm by enforcing constraints on their instance pairs.
They can be hard or soft, allowing different degrees of constraint violation.
Metric-based constraints map the constraints to the distance metric or similarity measure of the clustering algorithm.
Pairwise constraints are limited to partitional clustering.
So-called \ac{mlb} constraints introduce prior knowledge into hierarchical clustering.
A \ac{mlb} constraint 
\begin{eqnarray}
    MLB_{x y z} = (d_x, d_y, d_z)
    \label{eq:mlb_xyz}
\end{eqnarray}
specifies an agglomerating order on three data points $d_x, d_y, d_z$.
Symmetry applies for the first two elements.
For the \ac{mlb} constraint in \autoref{eq:mlb_xyz} both \autoref{def:mlb_implication} and \autoref{def:mlb_counterexample} must hold true.
This means that the must-link constraint between the first and second item is stronger than that between the first and third item.
\begin{definition}[\ac{mlb} implication]
     \label{def:mlb_implication}
    Given \ac{mlb} constraint $MLB_{x y z}$:
    $\forall c \in H: d_x \in c \wedge d_z \in c \rightarrow d_y \in c$
\end{definition}
\begin{definition}[\ac{mlb} counter example]
    \label{def:mlb_counterexample}
    Given \ac{mlb} constraint $MLB_{x y z}$:
    $\exists c \in H: d_x \in c \wedge d_y \in c \wedge d_z \notin c$
\end{definition}
A \ac{mlb} constraint for hierarchical clustering can be transformed into multiple pairwise constraints for partitional clustering by adding one must-link constraint between $d_x$ and $d_y$ and two cannot-links between $d_x, d_z$ and $d_y, d_z$~\cite{bade_hierarchical_2014}.


\subsubsection{\acs{hac} and \acs{ihac}}
The \ac{hac} algorithm iteratively merges the two most similar clusters into one.
The \ac{ihac} extends the \ac{hac} algorithm by not only considering similarity but also constraint violations~\cite{bade_hierarchical_2014}.

\textcolor{red}{was heißt ihac ist besser als hac? Stefan}


\subsubsection{Attribute Exploration}
\textcolor{red}{What is Attribute Exploration? \cite{ganter_conceptual_2016}}


\subsubsection{Incomplete knowledge}
\textcolor{red}{\ac{fca} with incomplete knowledge? \cite{holzer_incompleteFCA_2001}}

% copied from last paper
\input{chapters/extras/FCA}

\input{chapters/extras/fca_clusters}
\fcaclusterconnection{}
