\subsection{Gold-label Topic Hierarchy}
\label{sec:gt_mlb}

The gold-label-based method, consists of topic constraint generation, topic–document constraint mapping, and formal context construction.
\subsubsection{\ac{mlb} Constraints}

Starting from the three elements $\{x,y,z\}$ arranged as displayed in \autoref{fig:concept_lattice2mlb}, the mapping $\mathrm{MLB}$ produces the constraint $(x,y,z)$.

\latticeToMLB{}


\subsection{\acl{ae}\textcolor{red}{: Dominik}}

Since we deal with incomplete knowledge, we employ the an extension to traditional \ac{ae} as described in \autoref{subsec:basic_ae}, namely, Object Exploration in partial contexts, as described in \autoref{subsec:partial_cxt_ae}.

\todo{Widerspruchsfreiheit der Constraints!}\\
\todo{Incorporating MLB constraints during object exploration (von Stefan)}
\subsubsection{Object Exploration in Partial Contexts}
\label{subsec:partial_cxt_ae}

Classic Attribute Exploration, as defined by Ganter \cite{Ganter2016}, determines the implicational theory of a domain by interacting with a domain expert. Given a formal context $\mathbb{K} = (G, M, I)$, an attribute implication $A \to B$ (where $A, B \subseteq M$) holds if $A' \subseteq B'$.
However, since our goal is to find dependencies between data points based on their cluster membership, we employ \textbf{Object Exploration}. Due to the duality of FCA, Object Exploration on $\mathbb{K}$ is equivalent to Attribute Exploration on the transposed context $\mathbb{K}^T = (M, G, I^T)$.

An \textit{object implication} is of the form $X \to Y$ where $X, Y \subseteq G$ are sets of objects (here: data points). It is valid in $\mathbb{K}$ if every attribute (here: cluster $c \in M$) that contains all objects in $X$ also contains all objects in $Y$:
\begin{equation}
    \forall c \in M: X \subseteq c' \implies Y \subseteq c'
\end{equation}

In real-world clustering scenarios, the full hierarchy is not known \textit{a priori}. We therefore operate on a \textbf{Partial Context} (or incomplete context) $\mathbb{K}_{part} = (G, M, \{ \times, -, ? \}, J)$, following the theory of Burmeister and Holzer \cite{Holzer2001,BurmeisterHolzer}.
The incidence relation $J$ maps pairs $(g,m)$ to one of three values:
\begin{itemize}
    \item $\times$ (True): Object $g$ is definitely in cluster $m$.
    \item $-$ (False): Object $g$ is definitely not in cluster $m$.
    \item $?$ (Unknown): The relation is not yet determined.
\end{itemize}
The goal of the exploration is to replace the question marks with definite values ($\times$ or $-$) by querying "expert knowledge", in our case, the MLB constraints.

\todo{\cite{ferre_usability_2009}}

% \attrEx{}
\subsection{MLB Constraints as Exploratory Conditions}

We interpret an MLB constraint $MLB_{xyz} = (d_x, d_y, d_z)$ not merely as a distance inequality ($dist(x,y) < dist(x,z)$), but as a composite logical condition governing the object implications in the partial context.
From an FCA perspective, $MLB_{xyz}$ imposes two simultaneous requirements on the lattice structure: a \textit{Connectivity Requirement} (Validity) and a \textit{Separation Requirement} (Counter-example).

\begin{definition}[MLB Connectivity via Implication]
    \label{def:mlb_imp}
    The condition that $d_y$ is strictly closer to $d_x$ than $d_z$ implies that no cluster can contain the distant pair $\{d_x, d_z\}$ without also containing the intermediate point $d_y$. This enforces the validity of the object implication:
    \begin{equation}
        \{d_x, d_z\} \to \{d_y\}
    \end{equation}
    In the partial context, if the exploration algorithm proposes this implication, the "expert" (the constraint system) must confirm it. This confirmation allows the derivation of new information: for any cluster $c$ where $J(d_x, c)=\times$ and $J(d_z, c)=\times$, the value $J(d_y, c)$ must be set to $\times$, resolving potential question marks.
\end{definition}

\begin{definition}[MLB Separation via Counter-Example]
    \label{def:mlb_sep}
    The strict inequality ($<$) in the MLB constraint implies that there exists a level of granularity where $d_x$ and $d_y$ are clustered together, but $d_z$ is excluded. This constitutes a structural \textbf{counter-example} to the inverse implication $\{d_x, d_y\} \to \{d_z\}$.
    Formally, the context must allow for (or be extended to include) a cluster attribute $c_{sep} \in M$ such that:
    \begin{equation}
        d_x \in c'_{sep} \land d_y \in c'_{sep} \land d_z \notin c'_{sep}
    \end{equation}
\end{definition}

If the exploration were to accept $\{d_x, d_y\} \to \{d_z\}$, it would imply that $d_z$ is inseparable from the pair $(d_x, d_y)$, violating the strict order of the MLB constraint. Thus, the existence of such a counter-example is necessary to model the constraint correctly.

\subsection{The Exploration Algorithm with MLB Constraints}

Based on these definitions, we adapt the exploration algorithm to construct the hierarchy. The process starts with a partial context containing the data points $G$ and an initial set of (potentially empty or theoretical) cluster attributes $M$ initialized with $?$.

The interaction loop proceeds as follows:

\begin{enumerate}
    \item \textbf{Hypothesis Generation:} The algorithm computes the canonical basis of object implications valid in the current partial context.

    \item \textbf{Implication Query:} For a hypothesized implication $A \to B$, the system checks against the set of MLB constraints.
          \begin{itemize}
              \item If the implication corresponds to the connectivity requirement of an active constraint (e.g., $\{d_x, d_z\} \to \{d_y\}$ for $MLB_{xyz}$), it is \textbf{confirmed}. The context is updated by inferring $\times$ entries. \textcolor{orange}{Rest is ?.}
              \item If the implication contradicts the separation requirement (e.g., $\{d_x, d_y\} \to \{d_z\}$), it is \textbf{rejected}.
          \end{itemize}

    \item \textbf{Counter-Example Generation:} Upon rejection, the algorithm requires a counter-example. The system generates a new cluster column (or identifies an existing one) representing the separation state defined in Definition \ref{def:mlb_sep}, explicitly setting $J(d_z, c_{sep}) = -$ while maintaining $J(d_x, c_{sep}) = \times$ and $J(d_y, c_{sep}) = \times$. \textcolor{orange}{Rest is ?.}
\end{enumerate}

This iterative process ensures that the resulting concept lattice, and thus the hierarchical clustering, is fully consistent with the topological constraints imposed by the MLB triples.

\subsection{The Solution Space: Partial Contexts and Dendrograms}
\label{sec:solution_space}

A fundamental question in constrained clustering is whether the resulting model covers all valid clustering possibilities or forces an arbitrary selection. In our approach, the output of the object exploration process is a partial context $\mathbb{K}_{part} = (G, M, \{\times, -, ?\}, J)$. We argue that this context acts as a compact representation of the entire solution space of valid dendrograms.
\textcolor{orange}{Mathematischer Beweis?}

\subsubsection{Consistent Completions as Valid Hierarchies}
In the theory of incomplete knowledge, a partial context does not define a single closure system but rather a set of possible worlds. A \textit{completion} of $\mathbb{K}_{part}$ is a binary context $\mathbb{K}_{comp} = (G, M, I)$ where every $?$ in $J$ is replaced by either $\times$ or $-$ in $I$, such that all implications derived from the confirmed MLB constraints remain valid.

Let $\mathcal{D}$ be the set of all possible dendrograms (strictly nested hierarchies of object sets) over $G$. We define a dendrogram $D \in \mathcal{D}$ to be \textit{admissible} with respect to a set of MLB constraints $\Sigma$ if for every $(d_x, d_y, d_z) \in \Sigma$, the ultra-metric condition induced by $D$ satisfies $dist_D(x, y) < dist_D(x, z)$.

\begin{proposition}
    The partial context $\mathbb{K}_{part}$ derived from the exploration represents the set of all admissible dendrograms in the following sense:
    \begin{enumerate}
        \item Every admissible dendrogram corresponds to a consistent completion of $\mathbb{K}_{part}$.
        \item The question marks in $\mathbb{K}_{part}$ represent the degrees of freedom remaining in the clustering process after satisfying all expert constraints.
    \end{enumerate}
\end{proposition}

Consequently, the partial context captures the necessary structural properties enforced by the constraints. If $J(g, m) = \times$, object $g$ \textit{must} be in cluster $m$ in every valid hierarchy. If $J(g, m) = -$, it is strictly forbidden. The remaining question marks indicate regions where the constraints do not enforce a specific topology, allowing for variation in the final clustering granularity.

\subsubsection{Lattice Representation and Closed World Assumption}
While $\mathbb{K}_{part}$ represents the potential solution space, a concrete concept lattice is required for visualization and further processing.
Standard hierarchical clustering algorithms implicitly operate under a parsimony principle: clusters are not formed unless evident from data or constraints. We map this to FCA via a \textit{Closed World Assumption} (CWA) on the partial context.

Let $\mathbb{K}_{min}$ be the pessimistic completion of $\mathbb{K}_{part}$ where all remaining unknown entries are treated as false:
\begin{equation}
    I_{min}(g, m) := \begin{cases}
        \times & \text{if } J(g, m) = \times                          \\
        -      & \text{otherwise (i.e., if } - \text{ or } ? \text{)}
    \end{cases}
\end{equation}

The concept lattice $\underline{\mathfrak{B}}(\mathbb{K}_{min})$ derived from this completion is the minimal structure satisfying all MLB connectivity requirements.
It is important to note that a dendrogram is mathematically a tree, whereas $\underline{\mathfrak{B}}(\mathbb{K}_{min})$ is a lattice. The lattice contains all necessary clusters (concept extents) mandated by the constraints. Any specific dendrogram (a conflict-free hierarchy) consistent with the constraints is strictly embeddable in this lattice. Specifically, the clusters of such a dendrogram form a sub-order of the concept lattice's meet-semilattice.

Thus, our FCA-based approach generalizes the notion of a dendrogram: instead of forcing a single tree structure where data is ambiguous, the concept lattice visualizes the overlapping possibilities and the structural implications of the constraints, collapsing into a standard dendrogram only when the constraints (or the CWA) eliminate all degrees of freedom.
\textcolor{orange}{Ist wirklich minimal?}
