\subsection{Consistency-Lattice}
\label{sec:integrated_lattice_computation}
% Subsequently, we present our algorithm for the efficient computation of this structure, utilizing an on-the-fly irreducibility check and an alternating closure composition.
To integrate the background knowledge from the \ac{mlb} constraints with the latent knowledge from the \ac{lda} topic model, we construct the intersection of their respective closure systems.
Let $\mathbb{K}_{TM} = (G, M_{TM}, I_{TM})$ be the formal context derived from the \ac{lda} topic model. 
This context induces a standard \ac{fca} closure operator $\phi_{TM}(X) = X''$.

We seek the closure system containing those sets of documents that are closed under \emph{both} the \ac{mlb} constraints and the topic model logic:
\begin{equation}
    \mathcal{C}_{\cap} = \mathcal{C}_{\Sigma} \cap \mathcal{C}_{TM} = \{ A \subseteq G \mid \phi_{\Sigma}(A) = A \land \phi_{TM}(A) = A \}
\end{equation}

\subsubsection{Alternating Closure Composition}
The closure operator $\phi_\cap$ corresponding to this intersection is the limit of the alternating composition of $\phi_\Sigma$ and $\phi_{TM}$. For any set $X \subseteq G$, we define a sequence $X_0 = X$, $X_{k+1} = \phi_{TM}(\phi_\Sigma(X_k))$. Since $G$ is finite and closure operators are extensive and monotone, this sequence converges to a fixpoint: $\phi_\cap(X) = \lim_{k\to\infty} (\phi_{TM} \circ \phi_\Sigma)^k(X)$. This combined operator ensures that any generated concept is valid with respect to the expert constraints while also being statistically supported by the topic model.

\subsubsection{On-the-fly Computation of Irreducible Concepts}
Generating the full lattice for large datasets is computationally expensive. This challenge is further exacerbated by the preceding saturation step, which significantly increases the number of implications and thus the density of the closure system. However, to define a new, reduced formal context representing the merged knowledge, we only require the $\cap$-irreducible (meet-irreducible) elements of the lattice. An extent $A \in \mathcal{C}_\cap$ is $\cap$-irreducible if it cannot be represented as the intersection of its strict supersets.

Instead of generating all closed sets and filtering them post-hoc, we employ an optimized version of Ganter’s Next-Closure algorithm. During the canonical generation of closed sets, for a current closed set $A$, we perform an \emph{on-the-fly irreducibility check}. We inspect the immediate upper neighbors by tentatively adding elements $g \in G \setminus A$:

\begin{equation}
I_{sup}(A) = \bigcap_{g \in G \setminus A} \phi_\cap(A \cup \{g\})
\end{equation}

If $I_{sup}(A) = A$, then $A$ is the intersection of its upper neighbors and thus \textbf{reducible}. If $I_{sup}(A) \supsetneq A$, then $A$ is \textbf{irreducible}. 

This approach is crucial to our methodology: while constraint saturation ensures structural completeness, the on-the-fly irreducibility check effectively manages the resulting complexity. It discards redundant combinations immediately upon generation, leaving only the essential, structurally significant clusters. The resulting set of irreducible extents $\{E_1, \dots, E_k\}$ forms the attribute set of the final, knowledge-integrated formal context:
$$\mathbb{K}_{merged} = (G, \{m_1, \dots, m_k\}, I_{merged}),$$ where $g \, I_{merged} \, m_i \iff g \in E_i$.
We denote the concept lattice derived from $\mathbb{K}_{merged}$ as the \textit{consistency}-lattice, as it represents the hierarchy of concepts that are consistent with both the expert and data-driven knowledge.