\section{Method}
\label{sec:method}
% defines commands used below to insert datastructure (for better readability)
\input{chapters/extras/data_structure}

\subsection{\acs{mlb} constraints and \acs{fca}}

Implications are usually defined on attribute sets rather than on object sets \cite{fca-book}, but we will work with object exploration since it aligns directly with constrained hierarchical clustering.
\textcolor{gray}{The methodology proposed is directly applicable to both versions, differing only by the necessity of transposing the context for visualization via concept lattices.}
We denote the data points $\mathcal{D}=\{d_A, d_B, d_C, d_D, \dots\}$ the set of objects $G$ and the clusters $c_i \in \mathcal{C}$ the set of attributes $M$.
For simplicity, we explain our idea on the toy example given in \autoref{tab:formal_context}.
\formalcontexttable{}

The corresponding hierarchical cluster is given as a dendrogram in \autoref{fig:dendrogram}.
The bottom elements are data points in $\mathcal{D}$.
Iterative merging of clusters are indicated by agglomerating lines from the bottom to the top.
\dendogram{}

There are two directions to consider for proving the equivalence of both presentations.
First, we will outline how a hierarchical cluster can be turned into its corresponding formal context and next, we will describe the other direction.
In the following, we will describe the hierarchical cluster as a dendrogram.

\subsubsection{Hierarchical cluster to formal context.}
For each cluster on the lowest hierarchical level, i.e., the lowest line between leaf elements, add an attribute $c_i$ with $i > |\mathcal{C}|$.
Next, update the incidence relation $I = I \cup \{ d | d \in c_i\}$, i.e., add each data point $d$ of that cluster to the new attribute's extent, such that $(c_i)'$ contains all data points of that cluster.
Once all lower clusters have been processed, proceed with the next higher cluster hierarchy.
For each cluster emerging from sub-clusters $c_i, c_j \in \mathcal{C}$, add a new attribute $c_i \vee c_j$.
This attribute's extent contains all attributes in either one of the sub-cluster's extents, i.e., $(c_i \vee c_j)' = (c_i)'\cup(c_j)'$.
Proceed likewise for all clusters on that hierarchy and repeat the process for higher level emerging clusters.


\subsubsection{Formal context to hierarchical cluster.}
Given a formal context $(G,M,I)$ where $G$ denotes the set of data points $\mathcal{D}$ and $M$ denotes the set of clusters $c_i \in \mathcal{C}$, all objects $d \in G$ are leaf nodes of the dendogram.
For each attribute $a_i \in M$, identify any other attributes $a_j \in M$ whose extent is a superset of $a_i$'s extent.
Save those attributes in a successor list (or better: set) $s_{a_i}$.
% one could omit transitive constraints, but not necessary
To keep track of already processed cluster attributes, initialize a list $processed$.
To keep track of the constraints add a list of resulting constraints $r$.
%
Next, select one attribute $a_i$ with minimal extent cardinality $|(a_i)'|$, i.e., the smallest cluster.
For any pair of objects in $a_i$'s extent, i.e. data points $d_i, d_j \in (a_i)'$, add a pairwise \texttt{must-link} constraint to $r$.
To capture the hierarchical relation between the clusters, for $d_i, d_j$ and any object $d_k$ in $a_i$'s successor list's items' extent excluding $d_i, d_j$, i.e., data point $d_k \in (a)'\smallsetminus \{d_i, d_j\}$ for attribute $a \in s_{a_i}$, add a \texttt{\ac{mlb}} constraint $MLB_{d_i, d_j, d_k}$.
Once all \ac{mlb} constraints for the pair $d_i, d_j$ and objects $d_k \in (a)'\smallsetminus \{d_i, d_j\}$ have been added, the next pair in $(a_i)'$ can be processed.
Once all pairs in $(a_i)'$ have been processed, $a_i$ is added to $processed$.
Repeat the steps for the next attribute $a_j$, i.e., cluster, with minimal extent cardinality $|(a_j)'|$ which has not yet been processed, i.e., is not element of $processed$ is selected.
The algorithm ends once all cluster attributes have been added to $processed$.
%
The resulting set of \ac{mlb} constraints is models the solution space of possible clusters.
Any clustering algorithm, i.e., including \ac{hac} and \ac{ihac}, that respects these \textit{hard} constraints can produce a possible solution that lives within this solution space.

\subsubsection{Formal context to concept lattice.}
Given a formal context $(G,M,I)$ where $G$ denotes the set of data points $\mathcal{D}$ and $M$ denotes the set of clusters $c_i \in \mathcal{C}$, we can compute a meaningful concept lattice representation.
The resulting concept lattice for the formal context in \autoref{tab:formal_context} is given in \autoref{fig:concept_lattice}.

\lattice{}


\subsection{Deriving \ac{mlb} constraints from categories}

\textcolor{red}{BankSearch:}
Hierarchies occur naturally in data, often in the form of structural organisation in a file system or by personal topic hierarchies~\cite{bade_creating_2008}.
\ac{mlb} constraints can be derived as illustrated in Figure~\ref{fig:derive_mlb_constraints}.

\deriveMLBconstraints{}

\subsection{External Knowledge}

Besides existing file structures, we employed external knowledge bases such as WordNet~\cite{fellbaum2010wordnet} and topic models~\cite{gutekunst2025conceptual} for enriching the dataset with categories.

\subsubsection{WordNet}
Synsets etc.~\cite{fellbaum2010wordnet}
WordNet can be used to derive constraints from expert ontologies by mapping each category term (lemma) to the synsets (senses) in which it appears and then exploiting ``IS-A'' relations between synsets: terms that share a synset can yield synonymy-based must-link constraints (cf. Figure~\ref{fig:wordnet-must-link-constraints}), and terms whose synsets are connected by an ``IS-A'' path can yield hierarchical constraints (cf. Figure~\ref{fig:wordnet-mlb-constraints}) whose strength depends on the shortest path distance to a common ancestor.
\begin{figure}[t]
    \centering
    \begin{tikzpicture}[
            term/.style={draw, rounded corners, align=center, inner sep=3pt},
            rel/.style={->, >=latex}
        ]
        % Same synset -> must-link constraint
        \node[term] (a) {term A};
        \node[term, below=6mm of a] (b) {term B};
        \node[term, right=14mm of a] (s1) {synset S1\\(same sense)};
        \node[align=center, right=12mm of s1] {must-link};
        \draw[rel] (a) -- (s1);
        \draw[rel] (b) -- (s1);

    \end{tikzpicture}
    \caption{Deriving constraints from shared synsets in WordNet.}
    \label{fig:wordnet-must-link-constraints}
\end{figure}
\begin{figure}[t]
    \centering
    \begin{tikzpicture}[
            term/.style={draw, rounded corners, align=center, inner sep=3pt},
            rel/.style={->, >=latex}
        ]

        % Connected synsets -> hierarchical constraint
        \begin{scope}[xshift=80mm]
            \node[term] (c) {term C};
            \node[term, below=8mm of c] (d) {term D};
            \node[term, right=14mm of c] (s2) {synset S2};
            \node[term, below=8mm of s2] (s3) {synset S3};
            \node[term, right=14mm of s2] (anc) {common ancestor\\synset A};
            \node[align=center, right=12mm of anc] {hierarchical\\constraint\\(distance)};
            \draw[rel] (c) -- (s2);
            \draw[rel] (d) -- (s3);
            \draw[rel] (s2) -- node[above]{IS-A} (anc);
            \draw[rel] (s3) -- node[below]{IS-A} (anc);
        \end{scope}
    \end{tikzpicture}
    \caption{Deriving constraints from connected ``IS-A'' chains in WordNet.}
    \label{fig:wordnet-mlb-constraints}
\end{figure}
This mapping also provides a coverage signal, because terms absent from any synset cannot be constrained through WordNet.
In our analysis, coverage is high, and if ``groups'' is empty, there are no overlapping synsets among category terms.
Although about 3000 shared ``IS-A'' chains exist, only 22 term pairs have distance 1, so the strongest hierarchical relations are sparse.
These observations highlight limitations: constraints are only obtainable for terms represented in WordNet; must-link constraints may be unavailable if categories do not share synsets; and ``IS-A''/\ac{mlb} constraints can be weak when common ancestors are distant, which reduces their practical discriminative power and may introduce noise in constraint strength estimation.

\subsubsection{Topic Models}

A topic model is a generative model for documents, where documents are mixtures of $K$ topics and topics are mixtures of words (i.e., multinomial distribution over a word vocabulary).
Topic models rely on the bag-of-words assumption, i.e., ignoring the order of words.
We employed gensim's \ac{lda}\footnote{\url{https://radimrehurek.com/gensim/models/ldamodel.html} (30.12.2025)}.
\ac{lda} is unsupervised and based on statistical bayesian topic models~\cite{alghamdi_survey_2015}.

We build 200 latent topics across the data corpus, and extract the 10 most dominant topics per document each represented by the 10 most probable words per topic.
The resulting structure is \texttt{\{document $d_1$: [topic $t_{11}$: [word $w_{t_{11}-1}$, \dots, word $w_{t_{11}-10}$], topic $t_{12}$ \dots], document $d_2$:\dots \}}.
The number of initial topics (i.e., 200) is chosen arbitrarily.


\section{Dominik}
\todo{Widerspruchsfreiheit der Constraints!}
\subsection{Object Exploration in Partial Contexts}

Classic Attribute Exploration, as defined by Ganter \cite{Ganter2016}, determines the implicational theory of a domain by interacting with a domain expert. Given a formal context $\mathbb{K} = (G, M, I)$, an attribute implication $A \to B$ (where $A, B \subseteq M$) holds if $A' \subseteq B'$.
However, since our goal is to find dependencies between data points based on their cluster membership, we employ \textbf{Object Exploration}. Due to the duality of FCA, Object Exploration on $\mathbb{K}$ is equivalent to Attribute Exploration on the transposed context $\mathbb{K}^T = (M, G, I^T)$.

An \textit{object implication} is of the form $X \to Y$ where $X, Y \subseteq G$ are sets of objects (here: data points). It is valid in $\mathbb{K}$ if every attribute (here: cluster $c \in M$) that contains all objects in $X$ also contains all objects in $Y$:
\begin{equation}
    \forall c \in M: X \subseteq c' \implies Y \subseteq c'
\end{equation}

In real-world clustering scenarios, the full hierarchy is not known \textit{a priori}. We therefore operate on a \textbf{Partial Context} (or incomplete context) $\mathbb{K}_{part} = (G, M, \{ \times, -, ? \}, J)$, following the theory of Burmeister and Holzer \cite{Holzer2001,BurmeisterHolzer}.
The incidence relation $J$ maps pairs $(g,m)$ to one of three values:
\begin{itemize}
    \item $\times$ (True): Object $g$ is definitely in cluster $m$.
    \item $-$ (False): Object $g$ is definitely not in cluster $m$.
    \item $?$ (Unknown): The relation is not yet determined.
\end{itemize}
The goal of the exploration is to replace the question marks with definite values ($\times$ or $-$) by querying "expert knowledge", in our case, the MLB constraints.

\subsection{MLB Constraints as Exploratory Conditions}

We interpret an MLB constraint $MLB_{xyz} = (d_x, d_y, d_z)$ not merely as a distance inequality ($dist(x,y) < dist(x,z)$), but as a composite logical condition governing the object implications in the partial context.
From an FCA perspective, $MLB_{xyz}$ imposes two simultaneous requirements on the lattice structure: a \textit{Connectivity Requirement} (Validity) and a \textit{Separation Requirement} (Counter-example).

\begin{definition}[MLB Connectivity via Implication]
    \label{def:mlb_imp}
    The condition that $d_y$ is strictly closer to $d_x$ than $d_z$ implies that no cluster can contain the distant pair $\{d_x, d_z\}$ without also containing the intermediate point $d_y$. This enforces the validity of the object implication:
    \begin{equation}
        \{d_x, d_z\} \to \{d_y\}
    \end{equation}
    In the partial context, if the exploration algorithm proposes this implication, the "expert" (the constraint system) must confirm it. This confirmation allows the derivation of new information: for any cluster $c$ where $J(d_x, c)=\times$ and $J(d_z, c)=\times$, the value $J(d_y, c)$ must be set to $\times$, resolving potential question marks.
\end{definition}

\begin{definition}[MLB Separation via Counter-Example]
    \label{def:mlb_sep}
    The strict inequality ($<$) in the MLB constraint implies that there exists a level of granularity where $d_x$ and $d_y$ are clustered together, but $d_z$ is excluded. This constitutes a structural \textbf{counter-example} to the inverse implication $\{d_x, d_y\} \to \{d_z\}$.
    Formally, the context must allow for (or be extended to include) a cluster attribute $c_{sep} \in M$ such that:
    \begin{equation}
        d_x \in c'_{sep} \land d_y \in c'_{sep} \land d_z \notin c'_{sep}
    \end{equation}
\end{definition}

If the exploration were to accept $\{d_x, d_y\} \to \{d_z\}$, it would imply that $d_z$ is inseparable from the pair $(d_x, d_y)$, violating the strict order of the MLB constraint. Thus, the existence of such a counter-example is necessary to model the constraint correctly.

\subsection{The Exploration Algorithm with MLB Constraints}

Based on these definitions, we adapt the exploration algorithm to construct the hierarchy. The process starts with a partial context containing the data points $G$ and an initial set of (potentially empty or theoretical) cluster attributes $M$ initialized with $?$.

The interaction loop proceeds as follows:

\begin{enumerate}
    \item \textbf{Hypothesis Generation:} The algorithm computes the canonical basis of object implications valid in the current partial context.

    \item \textbf{Implication Query:} For a hypothesized implication $A \to B$, the system checks against the set of MLB constraints.
          \begin{itemize}
              \item If the implication corresponds to the connectivity requirement of an active constraint (e.g., $\{d_x, d_z\} \to \{d_y\}$ for $MLB_{xyz}$), it is \textbf{confirmed}. The context is updated by inferring $\times$ entries.
              \item If the implication contradicts the separation requirement (e.g., $\{d_x, d_y\} \to \{d_z\}$), it is \textbf{rejected}.
          \end{itemize}

    \item \textbf{Counter-Example Generation:} Upon rejection, the algorithm requires a counter-example. The system generates a new cluster column (or identifies an existing one) representing the separation state defined in Definition \ref{def:mlb_sep}, explicitly setting $J(d_z, c_{sep}) = -$ while maintaining $J(d_x, c_{sep}) = \times$ and $J(d_y, c_{sep}) = \times$.
\end{enumerate}

This iterative process ensures that the resulting concept lattice, and thus the hierarchical clustering, is fully consistent with the topological constraints imposed by the MLB triples.

\subsection{The Solution Space: Partial Contexts and Dendrograms}
\label{sec:solution_space}

A fundamental question in constrained clustering is whether the resulting model covers all valid clustering possibilities or forces an arbitrary selection. In our approach, the output of the object exploration process is a partial context $\mathbb{K}_{part} = (G, M, \{\times, -, ?\}, J)$. We argue that this context acts as a compact representation of the entire solution space of valid dendrograms.

\subsubsection{Consistent Completions as Valid Hierarchies}
In the theory of incomplete knowledge, a partial context does not define a single closure system but rather a set of possible worlds. A \textit{completion} of $\mathbb{K}_{part}$ is a binary context $\mathbb{K}_{comp} = (G, M, I)$ where every $?$ in $J$ is replaced by either $\times$ or $-$ in $I$, such that all implications derived from the confirmed MLB constraints remain valid.

Let $\mathcal{D}$ be the set of all possible dendrograms (strictly nested hierarchies of object sets) over $G$. We define a dendrogram $D \in \mathcal{D}$ to be \textit{admissible} with respect to a set of MLB constraints $\Sigma$ if for every $(d_x, d_y, d_z) \in \Sigma$, the ultra-metric condition induced by $D$ satisfies $dist_D(x, y) < dist_D(x, z)$.

\begin{proposition}
    The partial context $\mathbb{K}_{part}$ derived from the exploration represents the set of all admissible dendrograms in the following sense:
    \begin{enumerate}
        \item Every admissible dendrogram corresponds to a consistent completion of $\mathbb{K}_{part}$.
        \item The question marks in $\mathbb{K}_{part}$ represent the degrees of freedom remaining in the clustering process after satisfying all expert constraints.
    \end{enumerate}
\end{proposition}

Consequently, the partial context captures the necessary structural properties enforced by the constraints. If $J(g, m) = \times$, object $g$ \textit{must} be in cluster $m$ in every valid hierarchy. If $J(g, m) = -$, it is strictly forbidden. The remaining question marks indicate regions where the constraints do not enforce a specific topology, allowing for variation in the final clustering granularity.

\subsubsection{Lattice Representation and Closed World Assumption}
While $\mathbb{K}_{part}$ represents the potential solution space, a concrete concept lattice is required for visualization and further processing.
Standard hierarchical clustering algorithms implicitly operate under a parsimony principle: clusters are not formed unless evident from data or constraints. We map this to FCA via a \textit{Closed World Assumption} (CWA) on the partial context.

Let $\mathbb{K}_{min}$ be the pessimistic completion of $\mathbb{K}_{part}$ where all remaining unknown entries are treated as false:
\begin{equation}
    I_{min}(g, m) := \begin{cases}
        \times & \text{if } J(g, m) = \times                          \\
        -      & \text{otherwise (i.e., if } - \text{ or } ? \text{)}
    \end{cases}
\end{equation}

The concept lattice $\underline{\mathfrak{B}}(\mathbb{K}_{min})$ derived from this completion is the minimal structure satisfying all MLB connectivity requirements.
It is important to note that a dendrogram is mathematically a tree, whereas $\underline{\mathfrak{B}}(\mathbb{K}_{min})$ is a lattice. The lattice contains all necessary clusters (concept extents) mandated by the constraints. Any specific dendrogram (a conflict-free hierarchy) consistent with the constraints is strictly embeddable in this lattice. Specifically, the clusters of such a dendrogram form a sub-order of the concept lattice's meet-semilattice.

Thus, our FCA-based approach generalizes the notion of a dendrogram: instead of forcing a single tree structure where data is ambiguous, the concept lattice visualizes the overlapping possibilities and the structural implications of the constraints, collapsing into a standard dendrogram only when the constraints (or the CWA) eliminate all degrees of freedom.

\subsection{Deriving Formal Contexts from Datasets}
\textcolor{red}{TODO: How do we get the input (G, M, I) from which Datasets?
    Each document is exactly mapped to one category / topics:
    \cite{bade_hierarchical_2014}
    \begin{itemize}
        \item BankSearch Dataset
        \item Reuters Dataset (old)
    \end{itemize}
    Each document is cannot been mapped exactly to one category (but multiple):
    \begin{itemize}
        \item Reuters Dataset (new)
    \end{itemize}
    Expert Ontology as reference:
    \begin{itemize}
        \item WordNet \cite{fellbaum2010wordnet}
    \end{itemize}
    Probably same considerations on research designs by using object attribute matrices consisting of documents and topics exists.
    \cite{ceulemans_detecting_2010}
}
