\section{Case Study}
\label{sec:case_study}

We apply our method to the BankSearch dataset (\autoref{sec:banksearch_data}), for which we map topic-level \ac{mlb} constraints to document-level constraints (\autoref{sec:topic_document_mapping}).
The resulting \ac{mlb}-constrained concept lattice is then compared with a topic model-based iceberg lattice (Section~\ref{sec:lattice_comparison} and \ref{sec:consistency_vs_coherence}), followed by a discussion of methodological limitations (\autoref{sec:case_study_limitations}).

\input{chapters/06_dataset}


\subsection{Topic-Document Constraint Mapping}
\label{sec:topic_document_mapping}

We extract \ac{mlb} constraints at the topic level and map them to document-level constraints using the topic assignments.
Since each document belongs to exactly one topic, documents sharing the same topic form an equivalence class. Instead of translating constraints for every document individually, we select one representative document per topic and map the topic-level \ac{mlb} constraints to these representatives. 
This reduces the number of constraints while preserving the structure induced by the topic hierarchy.
\textcolor{red}{TODO: WHY EVEN EXISTING? DELETE TOPICS W/O DOCS BEFORE. We omit \ac{mlb} constraints involving categories that have no associated documents.}
For example, if documents $d_x$, $d_y$, and $d_z$ belong to topics $x$, $y$, and $z$, respectively, the topic-level constraint $(x,y,z)$ translates to the document-level constraint $(d_x, d_y, d_z)$.
For the BankSearch dataset, this procedure yields $81$ \ac{mlb} constraints.

% \input{chapters/08_case_study/dendogram_effectiveness_measure.tex}

\input{chapters/08_case_study/lattice_comparison.tex}

\subsection{Limitations \textcolor{orange}{NEW SECTION}}
\label{sec:case_study_limitations}

% review: I did not see a good discussion of the issues brought about by choosing the human-labelled data as "gold" (ground?) truth, and the LDA labelling as the system to be evaluated. But since this is a probabilistic assignment, they are going to be huge. Using a single threshold to realise the relevance of a topic to a document is not going to cut it, since it introduces a new hyper-parameter, the threshold, that has to be explored.

Several limitations should be considered when interpreting the results.
First, the BankSearch taxonomy comprises only two hierarchy levels and incomplete topic coverage (Section~\ref{sec:banksearch_data}).
In addition, some documents are empty or contain only numerical content.
To preserve a consistent document universe across experiments, such documents were retained using fallback preprocessing strategies, potentially introducing noise into the learned topic distributions.

\textcolor{red}{TODO: too long}
Second, the topic model-based representation depends on modeling assumptions and hyperparameters.
The document-topic incidence relation is obtained by thresholding probabilistic \ac{lda} assignments, making the resulting formal context sensitive to the selected threshold $\delta$.
Likewise, fixing the number of latent topics to the number of BankSearch categories improves comparability, but may constrain the model's latent structure discovery.

Finally, \ac{lda} represents only one class of topic modeling approaches.
Alternative models, including neural and nonparametric topic models, may yield substantially different document-topic relations and concept lattices.
